---
title: "TODO"
author: "Doug Kelkhoff"
date: "2/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Improve `warn` argument input options

In addition to which source of arguments should be warned for, include the
option to explicitly specify which arguments generate warnings.

Possible warnings
*  by sourcee:  `'...'`, `'dots'`, `'call'`
* all aesthetics (`aes`) or no aesthetics (`-aes`)
* specific aesthetics (`color`) or specific aesthetic exceptions (`-color`)

```{r, eval = FALSE}
ggpack(..., warn = list('...', 'dots', 'call', aes, -aes, color, -color))
```

# Make ggpackets lazy

`ggpack` commands should store the arguments of the 'packaging' process and only
evaluate once the plot is being constructed. This would allow for interception
and processing of arguments specified outside the ggpack calls. 

The following functionality is enabled or made more useful by lazy evaluation of
the ggpack calls:

## Remapping of aesthetics

Allow a `ggpack` call to rearrange aesthetics to reduce duplication of aesthetic
inputs and minimize user prefixes to handle aesthetic propegation.

With lazy evaluation of `ggpack` calls, this can even draw from aesthetics
mapped using `aes()` prior to the `ggpack` call

```{r, eval = FALSE}
# want use to be able to specify a single aesthetic, color, that affects the
# line and ribbon geoms, but allows fill to still be used for label backgrounds.
ggpack(geom_line, ...) + 
ggpack(geom_ribbon, remap = list(fill = color), ..., color = NA) + 
ggpack(geom_label, ...)
```

```{r}
# alternative syntax using ggplot-style 'reserved' symbols like ..density..
ggpack(geom_ribbon, fill = ..color.., ..., color = NA)
```