--- 
title: "Composing Templates & Functions" 
author: "Doug Kelkhoff" 
date: "`r Sys.time()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(
  echo = TRUE, 
  fig.align='center', 
  fig.width = 6,
  fig.height = 4,
  out.width = '600px', 
  out.height = '400px') 
library(ggpackets)
```

# Overview

`ggpackets` is most powerful when it's being used to build template functions. 
By using it as a framework for packaging components of a ggplot into a single 
object, it provides a simple framework for constructing full featured ggplot
wrappers. Most importantly, this will handle a wealth of ggplot-naitive
parameters, exposing complete control over the components of a plot with minimal
added syntax. That's not to say there is no work to be done - to elegantly
handle parameters intuitively and with minimal syntactic burden on the end user,
there might be some amount of argument handling in order to satisfy the ideal
user interface. The goal is to allow the programmer to focus on building around
the usability of their function instead of the nuances of constructing plots and handling interactivity with the extensible ggplot ecosystem.

# Composing a template

First, before we get too far along, `ggpackets` is nearly useless without
components from `ggplot2`. Let's get that loaded.

```{r libraries, message = FALSE, warning = FALSE}
library(ggplot2)
```

To build our first template, we'll simply add a few layers to a `ggpacket` 
object. We'll specify some aesthetic parameters - the size the line layer (`geom_line()`) and the point layer (`geom_point()`).

```{r simple_ggpacket}
ggpk_my_template <- ggpacket() + 
  geom_line(size = 1) + 
  geom_point(size = 3)
```

```{r simple_ggpacket_output}
ggplot(Loblolly) + 
  aes(x = age, y = height, color = Seed) + 
  ggpk_my_template + 
  ggtitle('Growth of Loblolly Pines')
```

Although this might be helpful for making quick, reusable code, it does not 
provide much flexibility and is for all intents and purposes equivalent to 
`list(geom_line(size = 1), geom_point(size = 3))`. Perhaps the syntax is a
little more familiar to those that use ggplot regularly, but aside from that,
this simplist form provides very little real value.

# Wrapping a template in a function

Where the power of `ggpackets` begins to shine is in its transparent argument 
handling. If we want to bundle up our function above, but also want to provide 
optional handling of additional ggplot parameters, then ggpackets provides much 
of that functionality with very little additional syntax. Here we use the
`ggpack()` function to articulate parameters to each layer and lump in any extra
paramaters passed to our function.

```{r function_ggpacket}
ggpk_my_template <- function(...) {
  ggpack(geom_line, size = 1, ...) +
  ggpack(geom_point, size = 3, ...)
}
```

```{r function_ggpacket_output}
ggplot(Loblolly) + 
  aes(x = age, y = height) + 
  ggpk_my_template(stat = 'summary', color = age, fun.data = mean_se, size = 5, alpha = 0.5) + 
  ggtitle('Average Growth of Loblolly Pines')
```

In the function we declared above, all arguments passed to our wrapper function
get filter down into all the internal layers. When we pass through the `...`, we
pass through any unnamed arguments from the parent function.

In fact, when we call the function, we can pass any arguments that would
typically get passed to either of the ggplot functions wrapped inside of it,
`geom_line` and `geom_point`. If we choose to pass both the `data` and `mapping`
parameters (typically set using the `ggplot()` and `aes()` functions when
constructing a ggplot object), we could even use this as a standalone function
as shown below.

> Be aware that this usage isn't particularly intuitive without a bit of added
documentation. If you're building a function for someone else to use or to
export from a package you're making, it's probably worth writing up some good
documentation about what paramters you expect people to use.

```{r function_ggpacket_standalone_output}
ggpk_my_template(Loblolly, x = age, y = height, color = Seed)
```


# Granular Parameter Passing

There is one lingering issue. In this case, all parameters are being passed to 
both of the ggplot layers wrapped inside of our function. You can see the impact
of this issue above, because both the size (thickness) of the lines and the size
(diameter) of the points changed when we set a new size. If we want to control 
one or the other independently, we need a more precise way of passing
parameters. This can be addressed by adding an identifying string to each of our
layers using the `id` parameter to `ggpack()`.

If we want to restrict parameters to only one layer or another, `ggpackets` can 
handle the filtering of parameters through to each of the layers by parsing the
parameters for prefixes that match that particular layer's `id`. For example, if
a layer has an `id = 'my_ggplot_layer'`, then we might pass the function a
parameter like `my_ggplot_layer.size` (the id and parameter name separated by a
period) to set the size parameter of that particular layer without affecting the
others.

```{r granular_ggpacket}
ggpk_my_template <- function(...) {
  ggpack(geom_line, id = 'line', size = 1, ...) + 
  ggpack(geom_point, id = 'point', size = 3, ...)
}
```

```{r granular_ggpacket_output}
ggplot(Loblolly) + 
  aes(x = age, y = height, color = Seed) + 
  ggpk_my_template(point.size = 5, line.size = 0.5) + 
  ggtitle('Average Growth of Loblolly Pines')
```

Lists are also permitted as an `id`, allowing for any one of a number of
different prefixes to capture arguments for a particular layer.

```{r granular_ggpacket2}
ggpk_my_template <- function(...) {
  ggpack(geom_line, id = c('all', 'line'), size = 1, ...) + 
  ggpack(geom_point, id = c('all', 'point'), size = 3, ...)
}
```

```{r granular_ggpacket_output2}
ggplot(Loblolly) + 
  aes(x = age, y = height) + 
  ggpk_my_template(all.stat = 'summary', all.fun.data = mean_se,
    point.size = 5, line.size = 0.5) + 
  ggtitle('Average Growth of Loblolly Pines')
```

# Default, User and Fixed Parameters

The ordering of the parameters passed to any layer carries its own significance
and can be used to exercise further control over how your function operates. If
a parameter is passed to the layer more than once, the last instance always
takes priority. This includes parameters passed as a list via the `dots`
parameter. This allows calls to be constructed in a way that dictates which
arguments are set in stone and which can be ovewritten by the user.

Take for example the following function where all ellipses parameters are passed
through to the layers using the `dots` parameter. In the first layer, `size` is
declared before the parameters passed in by the user, whereas the second has
`size` passed after the user parameters.

```{r parameters_ggpacket}
ggpk_my_template <- function(...) {
  ggpack(geom_line, id = 'line', size = 3, ...) + 
  ggpack(geom_point, id = 'point', ..., size = 3)
}
```

The user will be able to set the size of the line layer, but will be unable to 
change the size of the point layer because the last `size` parameter will always
take precedent (even if it's passed with a prefix, like `point.size`). This
gives considerable control over what impact the user can have over the plot
template.

```{r parameters_ggpacket_output}
ggplot(Loblolly) + 
  aes(x = age, y = height, color = Seed) + 
  ggpk_my_template(line.size = 0.1, point.size = 10) + 
  ggtitle('Average Growth of Loblolly Pines')
```